---
title: "methClust applications (0/1 scale) - Birds data"
author: "Kushal K Dey"
date: "12/2/2017"
output: html_document
---

**methClust** is designed as a generic Binomial Grade of Membership model implementing software. A specific setting includes when the data is Bernoulli - that is observed versus unoberved. An example of such a setting is the presence-absence matrix from the bird abundance data.

## Loading Birds Abundance Data

```{r warning = FALSE, message = FALSE}
library(ecostructure)
library(methClust)
```

```{r}
data <- get(load(system.file("extdata", "HimalayanBirdsData.rda",
                             package = "ecostructure")))
taxonomic_counts <- t(exprs(data))
```

```{r}
source("../../methClust/R/meth_topics.R")
source("../../methClust/R/meth_tpx.R")
source("../../methClust/R/meth_count.R")
library(slam)
```

## Presence absence Data

```{r}
presence_absence <- taxonomic_counts
presence_absence[presence_absence > 0] = 1
absence_presence <- 1 - presence_absence
```

## Application of methClust

apply **methClust** with the `presence_absence` matrix as analog to the methylation matrix and `absence_presence` matrix as analog to the unmethylation matrix.

```{r eval=TRUE, warning = FALSE, message = FALSE}
topics_list <- list()

for(n in 1:100){
set.seed(100+n)
topics_list[[n]] <- meth_topics(presence_absence, absence_presence, 
                      K=2, tol = 1, use_squarem = FALSE, NUM_INDICES_START = 10)
}
```

## Visualization of methClust

```{r}
ind_max <- which.max(unlist(lapply(topics_list, function(x) return(x$L))))
topics <- topics_list[[ind_max]]
grid_metadata <- pData(phenoData(data))
head(grid_metadata)
elevation_metadata=grid_metadata$Elevation;
east_west_dir = grid_metadata$WorE;
omega <- topics$omega
BlockStructure(omega, blocker_metadata = east_west_dir,
               order_metadata = elevation_metadata,
               yaxis_label = "Elevation",
               levels_decreasing = FALSE)
```


```{r}
topics$L
```

```{r}
sessionInfo()
```

