---
title: "Global Map Motifs"
author: "Alex White"
date: "11/29/2017"
output: html_document
---

```{r setup, include=FALSE}
library(ecostructure)
library(Biobase)
library(rasterVis)
library(gtools)
library(bigalgebra)
library(sp)
library(rgdal)

```

## Global Breeding Bird - Regional Motifs

We compiled a dataset of all breeding bird species ranges to produce assemblage dispersion fields for each 1 x 1 degree map cell on earth. Of the total 64800 map cells covering the globe, we discarded cells that did not have at least 10 percent land cover (i.e., cells that covered ocean), with 17441 cells remaining after pruning the dataset. This step is important, as we do not want sea birds distributions to distort or overwhelm the terrestrial patterns that we are investigating using the GoM model.  

For each dispersion field (which includes the global distributions of all birds that are present within that cell), we vectorized the field and stacked the vectors within a single matix.  Each vector thus contained a count for each of the 64800 global 1x1 degree cells, and the resulting matrix was 17441 x 64800. Because many of the cells contained a count of 0 across all 17441 samples (over ocean), we dropped those columns from the matrix, resulting in a 17441 x 17808 matrix.  We ran the standard GoM model on the resulting counts matrix to identify global regional motifs. Given the results we observed for Himalayan birds, we hypothesized that the map motifs produced by the global analysis would be partitioned along a boundary of freezing temperatures (<0 degrees C mean January temp).  We find this hypothesis well supported by the resulting global map motifs. This is a compelling result, as birds are homeothermic and therfore have no strict limit imposed by freezing.  We suggest the influence of this climatic boundary instead reflects habitat affinities, as plant species, which are indeed strongly impacted by freezing temperatures, also turnover stongly accross the freezing line (this has been atributed to the trade off between frost tolerance vs. growth [competition]). We find that Himalayan tree species also turover at the freezing line in the Himalayas, supporting our assertion that the turnover in birds follows habitat changes.  

First, I plot the freezing boundary around the globe. 
```{r, echo=F, message=FALSE, warning=FALSE}

crs <- CRS("+proj=longlat +datum=WGS84")

myPalette <- colorRampPalette(rev(c("red","lemonchiffon","blue")))
climate <- getData('worldclim', var='bio', res=10)

temperate <- climate$bio6
tropical <- climate$bio6

temperate[temperate > 0] <- NA
tropical[tropical < 0] <- NA

ocean <- readOGR(dsn = path.expand("../data/ne_110m_ocean/ne_110m_ocean.shp"),
                 layer ="ne_110m_ocean")
                 
coast <- readOGR(dsn = path.expand("../data/ne_110m_coastline/ne_110m_coastline.shp"),
                 layer ="ne_110m_coastline")
                 
myTheme <- rasterTheme(region = myPalette(100))
levelplot(climate$bio6, 
          contour=F,
          margin=F,
          main=paste("Min Temp Coldest Month (white cells are < -25°C)"),
          par.settings = myTheme,
          at=seq(-250, 250, length.out=100),
          colorkey=list(at=seq(0, 100, 1),
                        labels=list(at=c(0,25,50,75,100),
                                    labels=c("-25°C","-12.5°C","0°C","12.5°C", "25°C")))) +
  latticeExtra::layer(sp.lines(coast, col='black', lwd=0.5))+
  latticeExtra::layer(sp.polygons(ocean, fill='grey', col='black'))


plot(tropical,col=myPalette(100),main="Min Temp Coldest Month, Above Freezing",zlim=c(-500,500))
plot(coast, add = T)
plot(temperate,col=myPalette(100),main="Min Temp Coldest Month, Below Freezing",zlim=c(-500,500))
plot(coast, add = T)

```

## K = 2

Results for K = 2:

```{r, echo=FALSE}
####
######## k2
#load(file="birds_cluster_2.rda")
out <- get(load(file="../output/birds_meth_cluster_10.rda"))
K10_global_topics <- out

topics_omega <- K10_global_topics$omega

proj <- CRS(' +proj=longlat +ellps=WGS84')
map.extent <- extent(-180,180,-90,90)

newtheme <- rasterTheme(region = colorRampPalette(c("darkseagreen3","orange","red"))( 1000 ))

v <- rep(0,64800)
v[as.numeric(rownames(topics_omega))] = topics_omega[,1]
motif_1_K2 <- matrix(v, nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_omega))] = topics_theta[,2]
motif_2_K2 <- matrix(v,nrow = 180, ncol = 360)


xmin <- -180
xmax <- 180
ymin <- -90
ymax <- 90

motif_K2_maps <- list()

motif_K2_maps[[1]] <- raster(motif_1_K2, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K2_maps[[2]] <- raster(motif_2_K2, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)



e <- extent(-180, 180, -60, 90)

motif.K2.theta.maps <- list()
for (i in 1:length(motif_K2_maps)){
  motif.K2.theta.maps[[i]] <- levelplot(crop(motif_K2_maps[[i]],e), 
                                        main=paste("Map cell contributions to Motif ",i, ", K = 2"),
                                        par.settings=newtheme, 
                                        contour=F,margin=FALSE,
                                        colorkey=list(at=seq(0, 100, 1),
                                                      labels=list(at=c(0,100),
                                                                  labels=c("Low", "High"))))+
    latticeExtra::layer(sp.lines(coast, col='black', lwd=0.5))+
    latticeExtra::layer(sp.polygons(ocean, fill='grey', col='black'))
}

for(k in 1:10){
  motif.K2.theta.maps[[k]]
}

```

## K = 3

Results for K = 3:

```{r, echo=FALSE}

####
######## k3
load(file="birds_cluster_3.rda")
K3_global_topics <- out

topics_theta <- K3_global_topics$theta

proj <- CRS(' +proj=longlat +ellps=WGS84')
map.extent <- extent(-180,180,-90,90)

newtheme <- rasterTheme(region = colorRampPalette(c("darkseagreen3","orange","red"))( 1000 ))
#newtheme <- colorRampPalette(c("darkseagreen3","orange","red"))( 100 )


v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,1]
motif_1_K3 <- matrix(v, nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,2]
motif_2_K3 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,3]
motif_3_K3 <- matrix(v,nrow = 180, ncol = 360)



xmin <- -180
xmax <- 180
ymin <- -90
ymax <- 90

motif_K3_maps <- list()

motif_K3_maps[[1]] <- raster(motif_1_K3, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K3_maps[[2]] <- raster(motif_2_K3, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K3_maps[[3]] <- raster(motif_3_K3, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)


e <- extent(-180, 180, -60, 90)

motif.K3.theta.maps <- list()
for (i in 1:length(motif_K3_maps)){
  motif.K3.theta.maps[[i]] <- levelplot(crop(motif_K3_maps[[i]],e), 
                                        main=paste("Map cell contributions to Motif ",i, ", K = 3"),
                                        par.settings=newtheme, 
                                        contour=F,margin=FALSE,
                                        colorkey=list(at=seq(0, 100, 1),
                                                      labels=list(at=c(0,100),
                                                                  labels=c("Low", "High"))))+
    latticeExtra::layer(sp.lines(coast, col='black', lwd=0.5))+
    latticeExtra::layer(sp.polygons(ocean, fill='grey', col='black'))
}



```

## K = 4

Results for K = 4:

```{r, echo=FALSE}
load(file="birds_cluster_4.rda")
K4_global_topics <- out

topics_theta <- K4_global_topics$theta

proj <- CRS(' +proj=longlat +ellps=WGS84')
map.extent <- extent(-180,180,-90,90)

newtheme <- rasterTheme(region = colorRampPalette(c("darkseagreen3","orange","red"))( 1000 ))
#newtheme <- colorRampPalette(c("darkseagreen3","orange","red"))( 100 )


v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,1]
motif_1_K4 <- matrix(v, nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,2]
motif_2_K4 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,3]
motif_3_K4 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,4]
motif_4_K4 <- matrix(v,nrow = 180, ncol = 360)

xmin <- -180
xmax <- 180
ymin <- -90
ymax <- 90

motif_K4_maps <- list()

motif_K4_maps[[1]] <- raster(motif_1_K4, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K4_maps[[2]] <- raster(motif_2_K4, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K4_maps[[3]] <- raster(motif_3_K4, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K4_maps[[4]] <- raster(motif_4_K4, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)

e <- extent(-180, 180, -60, 90)

motif.K4.theta.maps <- list()
for (i in 1:length(motif_K4_maps)){
  motif.K4.theta.maps[[i]] <- levelplot(crop(motif_K4_maps[[i]],e), 
                                        main=paste("Map cell contributions to Motif ",i, ", K = 4"),
                                        par.settings=newtheme, 
                                        contour=F,margin=FALSE,
                                        colorkey=list(at=seq(0, 100, 1),
                                                      labels=list(at=c(0,100),
                                                                  labels=c("Low", "High"))))+
    latticeExtra::layer(sp.lines(coast, col='black', lwd=0.5))+
    latticeExtra::layer(sp.polygons(ocean, fill='grey', col='black'))
}

motif.K4.theta.maps[[1]]
motif.K4.theta.maps[[2]]
motif.K4.theta.maps[[3]]
motif.K4.theta.maps[[4]]

```

## K = 5

Results for K = 5:

```{r, echo=FALSE}
load(file="birds_cluster_5.rda")
K5_global_topics <- out

topics_theta <- K5_global_topics$theta

proj <- CRS(' +proj=longlat +ellps=WGS84')
map.extent <- extent(-180,180,-90,90)

newtheme <- rasterTheme(region = colorRampPalette(c("darkseagreen3","orange","red"))( 1000 ))
#newtheme <- colorRampPalette(c("darkseagreen3","orange","red"))( 100 )


v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,1]
motif_1_K5 <- matrix(v, nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,2]
motif_2_K5 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,3]
motif_3_K5 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,4]
motif_4_K5 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,5]
motif_5_K5 <- matrix(v,nrow = 180, ncol = 360)

xmin <- -180
xmax <- 180
ymin <- -90
ymax <- 90

motif_K5_maps <- list()

motif_K5_maps[[1]] <- raster(motif_1_K5, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K5_maps[[2]] <- raster(motif_2_K5, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K5_maps[[3]] <- raster(motif_3_K5, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K5_maps[[4]] <- raster(motif_4_K5, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K5_maps[[5]] <- raster(motif_5_K5, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)

e <- extent(-180, 180, -60, 90)

motif.K5.theta.maps <- list()
for (i in 1:length(motif_K5_maps)){
  motif.K5.theta.maps[[i]] <- levelplot(crop(motif_K5_maps[[i]],e), 
                                        main=paste("Map cell contributions to Motif ",i, ", K = 5"),
                                        par.settings=newtheme, 
                                        contour=F,margin=FALSE,
                                        colorkey=list(at=seq(0, 100, 1),
                                                      labels=list(at=c(0,100),
                                                                  labels=c("Low", "High"))))+
    latticeExtra::layer(sp.lines(coast, col='black', lwd=0.5))+
    latticeExtra::layer(sp.polygons(ocean, fill='grey', col='black'))
}

motif.K5.theta.maps[[1]]
motif.K5.theta.maps[[2]]
motif.K5.theta.maps[[3]]
motif.K5.theta.maps[[4]]
motif.K5.theta.maps[[5]]

```

## K = 6

Results for K = 6:

```{r, echo=FALSE}
load(file="birds_cluster_6.rda")
K6_global_topics <- out

topics_theta <- K6_global_topics$theta

proj <- CRS(' +proj=longlat +ellps=WGS84')
map.extent <- extent(-180,180,-90,90)

newtheme <- rasterTheme(region = colorRampPalette(c("darkseagreen3","orange","red"))( 1000 ))
#newtheme <- colorRampPalette(c("darkseagreen3","orange","red"))( 100 )

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,1]
motif_1_K6 <- matrix(v, nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,2]
motif_2_K6 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,3]
motif_3_K6 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,4]
motif_4_K6 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,5]
motif_5_K6 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,6]
motif_6_K6 <- matrix(v,nrow = 180, ncol = 360)

xmin <- -180
xmax <- 180
ymin <- -90
ymax <- 90

motif_K6_maps <- list()

motif_K6_maps[[1]] <- raster(motif_1_K6, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K6_maps[[2]] <- raster(motif_2_K6, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K6_maps[[3]] <- raster(motif_3_K6, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K6_maps[[4]] <- raster(motif_4_K6, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K6_maps[[5]] <- raster(motif_5_K6, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K6_maps[[6]] <- raster(motif_6_K6, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)

e <- extent(-180, 180, -60, 90)

motif.K6.theta.maps <- list()
for (i in 1:length(motif_K6_maps)){
  motif.K6.theta.maps[[i]] <- levelplot(crop(motif_K6_maps[[i]],e), 
                                        main=paste("Map cell contributions to Motif ",i, ", K = 6"),
                                        par.settings=newtheme, 
                                        contour=F,margin=FALSE,
                                        colorkey=list(at=seq(0, 100, 1),
                                                      labels=list(at=c(0,100),
                                                                  labels=c("Low", "High"))))+
    latticeExtra::layer(sp.lines(coast, col='black', lwd=0.5))+
    latticeExtra::layer(sp.polygons(ocean, fill='grey', col='black'))
}

motif.K6.theta.maps[[1]]
motif.K6.theta.maps[[2]]
motif.K6.theta.maps[[3]]
motif.K6.theta.maps[[4]]
motif.K6.theta.maps[[5]]
motif.K6.theta.maps[[6]]

```

## K = 7

Results for K = 7:

```{r, echo=FALSE}
load(file="birds_cluster_7.rda")
K7_global_topics <- out

topics_theta <- K7_global_topics$theta

proj <- CRS(' +proj=longlat +ellps=WGS84')
map.extent <- extent(-180,180,-90,90)

newtheme <- rasterTheme(region = colorRampPalette(c("darkseagreen3","orange","red"))( 1000 ))
#newtheme <- colorRampPalette(c("darkseagreen3","orange","red"))( 100 )

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,1]
motif_1_K7 <- matrix(v, nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,2]
motif_2_K7 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,3]
motif_3_K7 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,4]
motif_4_K7 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,5]
motif_5_K7 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,6]
motif_6_K7 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,7]
motif_7_K7 <- matrix(v,nrow = 180, ncol = 360)

xmin <- -180
xmax <- 180
ymin <- -90
ymax <- 90

motif_K7_maps <- list()

motif_K7_maps[[1]] <- raster(motif_1_K7, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K7_maps[[2]] <- raster(motif_2_K7, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K7_maps[[3]] <- raster(motif_3_K7, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K7_maps[[4]] <- raster(motif_4_K7, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K7_maps[[5]] <- raster(motif_5_K7, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K7_maps[[6]] <- raster(motif_6_K7, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K7_maps[[7]] <- raster(motif_7_K7, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)

e <- extent(-180, 180, -60, 90)

motif.K7.theta.maps <- list()
for (i in 1:length(motif_K7_maps)){
  motif.K7.theta.maps[[i]] <- levelplot(crop(motif_K7_maps[[i]],e), 
                                        main=paste("Map cell contributions to Motif ",i, ", K = 7"),
                                        par.settings=newtheme, 
                                        contour=F,margin=FALSE,
                                        colorkey=list(at=seq(0, 100, 1),
                                                      labels=list(at=c(0,100),
                                                                  labels=c("Low", "High"))))+
    latticeExtra::layer(sp.lines(coast, col='black', lwd=0.5))+
    latticeExtra::layer(sp.polygons(ocean, fill='grey', col='black'))
}

motif.K7.theta.maps[[1]]
motif.K7.theta.maps[[2]]
motif.K7.theta.maps[[3]]
motif.K7.theta.maps[[4]]
motif.K7.theta.maps[[5]]
motif.K7.theta.maps[[6]]
motif.K7.theta.maps[[7]]

```

## K = 8

Results for K = 8:

```{r, echo=FALSE}

load(file="birds_cluster_8.rda")
K8_global_topics <- out

topics_theta <- K8_global_topics$theta

proj <- CRS(' +proj=longlat +ellps=WGS84')
map.extent <- extent(-180,180,-90,90)

newtheme <- rasterTheme(region = colorRampPalette(c("darkseagreen3","orange","red"))( 1000 ))
#newtheme <- colorRampPalette(c("darkseagreen3","orange","red"))( 100 )

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,1]
motif_1_K8 <- matrix(v, nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,2]
motif_2_K8 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,3]
motif_3_K8 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,4]
motif_4_K8 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,5]
motif_5_K8 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,6]
motif_6_K8 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,7]
motif_7_K8 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,8]
motif_8_K8 <- matrix(v,nrow = 180, ncol = 360)

xmin <- -180
xmax <- 180
ymin <- -90
ymax <- 90

motif_K8_maps <- list()

motif_K8_maps[[1]] <- raster(motif_1_K8, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K8_maps[[2]] <- raster(motif_2_K8, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K8_maps[[3]] <- raster(motif_3_K8, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K8_maps[[4]] <- raster(motif_4_K8, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K8_maps[[5]] <- raster(motif_5_K8, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K8_maps[[6]] <- raster(motif_6_K8, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K8_maps[[7]] <- raster(motif_7_K8, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K8_maps[[8]] <- raster(motif_8_K8, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)

e <- extent(-180, 180, -60, 90)

motif.K8.theta.maps <- list()
for (i in 1:length(motif_K8_maps)){
  motif.K8.theta.maps[[i]] <- levelplot(crop(motif_K8_maps[[i]],e), 
                                        main=paste("Map cell contributions to Motif ",i, ", K = 8"),
                                        par.settings=newtheme, 
                                        contour=F,margin=FALSE,
                                        colorkey=list(at=seq(0, 100, 1),
                                                      labels=list(at=c(0,100),
                                                                  labels=c("Low", "High"))))+
    latticeExtra::layer(sp.lines(coast, col='black', lwd=0.5))+
    latticeExtra::layer(sp.polygons(ocean, fill='grey', col='black'))
}

motif.K8.theta.maps[[1]]
motif.K8.theta.maps[[2]]
motif.K8.theta.maps[[3]]
motif.K8.theta.maps[[4]]
motif.K8.theta.maps[[5]]
motif.K8.theta.maps[[6]]
motif.K8.theta.maps[[7]]
motif.K8.theta.maps[[8]]

```

## K = 9

Results for K = 9:

```{r, echo=FALSE}

load(file="birds_cluster_9.rda")
K9_global_topics <- out


topics_theta <- K9_global_topics$theta

proj <- CRS(' +proj=longlat +ellps=WGS84')
map.extent <- extent(-180,180,-90,90)

newtheme <- rasterTheme(region = colorRampPalette(c("darkseagreen3","orange","red"))( 1000 ))
#newtheme <- colorRampPalette(c("darkseagreen3","orange","red"))( 100 )

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,1]
motif_1_K9 <- matrix(v, nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,2]
motif_2_K9 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,3]
motif_3_K9 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,4]
motif_4_K9 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,5]
motif_5_K9 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,6]
motif_6_K9 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,7]
motif_7_K9 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,8]
motif_8_K9 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,9]
motif_9_K9 <- matrix(v,nrow = 180, ncol = 360)

xmin <- -180
xmax <- 180
ymin <- -90
ymax <- 90

motif_K9_maps <- list()

motif_K9_maps[[1]] <- raster(motif_1_K9, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K9_maps[[2]] <- raster(motif_2_K9, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K9_maps[[3]] <- raster(motif_3_K9, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K9_maps[[4]] <- raster(motif_4_K9, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K9_maps[[5]] <- raster(motif_5_K9, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K9_maps[[6]] <- raster(motif_6_K9, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K9_maps[[7]] <- raster(motif_7_K9, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K9_maps[[8]] <- raster(motif_8_K9, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K9_maps[[9]] <- raster(motif_9_K9, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)

e <- extent(-180, 180, -60, 90)

motif.K9.theta.maps <- list()
for (i in 1:length(motif_K9_maps)){
  motif.K9.theta.maps[[i]] <- levelplot(crop(motif_K9_maps[[i]],e),
                                        main=paste("Map cell contributions to Motif ",i, ", K = 9"),
                                        par.settings=newtheme, 
                                        contour=F,margin=F,
                                        colorkey=list(at=seq(0, 100, 1),
                                                      labels=list(at=c(0,100),
                                                                  labels=c("Low", "High"))))+
    latticeExtra::layer(sp.lines(coast, col='black', lwd=0.5))+
    latticeExtra::layer(sp.polygons(ocean, fill='grey', col='black'))
}

motif.K9.theta.maps[[1]]
motif.K9.theta.maps[[2]]
motif.K9.theta.maps[[3]]
motif.K9.theta.maps[[4]]
motif.K9.theta.maps[[5]]
motif.K9.theta.maps[[6]]
motif.K9.theta.maps[[7]]
motif.K9.theta.maps[[8]]
motif.K9.theta.maps[[9]]

```

## K = 10

Results for K = 10:

```{r, echo=FALSE}

######## k10
K10_global_topics <- get(load(file="../output/birds_meth_cluster_10.rda"))

topics_theta <- K10_global_topics$omega

proj <- CRS(' +proj=longlat +ellps=WGS84')
map.extent <- extent(-180,180,-90,90)

newtheme <- rasterTheme(region = colorRampPalette(c("darkseagreen3","orange","red"))( 1000 ))
#newtheme <- colorRampPalette(c("darkseagreen3","orange","red"))( 100 )

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,1]
motif_1_K10 <- matrix(v, nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,2]
motif_2_K10 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,3]
motif_3_K10 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,4]
motif_4_K10 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,5]
motif_5_K10 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,6]
motif_6_K10 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,7]
motif_7_K10 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,8]
motif_8_K10 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,9]
motif_9_K10 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,10]
motif_10_K10 <- matrix(v,nrow = 180, ncol = 360)


xmin <- -180
xmax <- 180
ymin <- -90
ymax <- 90

motif_K10_maps <- list()

motif_K10_maps[[1]] <- raster(motif_1_K10, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K10_maps[[2]] <- raster(motif_2_K10, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K10_maps[[3]] <- raster(motif_3_K10, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K10_maps[[4]] <- raster(motif_4_K10, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K10_maps[[5]] <- raster(motif_5_K10, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K10_maps[[6]] <- raster(motif_6_K10, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K10_maps[[7]] <- raster(motif_7_K10, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K10_maps[[8]] <- raster(motif_8_K10, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K10_maps[[9]] <- raster(motif_9_K10, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K10_maps[[10]] <- raster(motif_10_K10, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)


e <- extent(-180, 180, -60, 90)

motif.K10.theta.maps <- list()
for (i in 1:length(motif_K10_maps)){
  motif.K10.theta.maps[[i]] <- levelplot(crop(motif_K10_maps[[i]],e),
                                        main=paste("Map cell contributions to Motif ",i, ", K = 10"),
                                        par.settings=newtheme, 
                                        contour=F,margin=F,
                                        colorkey=list(at=seq(0, 100, 1),
                                                      labels=list(at=c(0,100),
                                                                  labels=c("Low", "High"))))+
    latticeExtra::layer(sp.lines(coast, col='black', lwd=0.5))+
    latticeExtra::layer(sp.polygons(ocean, fill='grey', col='black'))
}

motif.K10.theta.maps[[1]]
motif.K10.theta.maps[[2]]
motif.K10.theta.maps[[3]]
motif.K10.theta.maps[[4]]
motif.K10.theta.maps[[5]]
motif.K10.theta.maps[[6]]
motif.K10.theta.maps[[7]]
motif.K10.theta.maps[[8]]
motif.K10.theta.maps[[9]]
motif.K10.theta.maps[[10]]

```

## K = 12

Results for K = 12:
```{r, echo=FALSE}
######## k12
load(file="birds_cluster_12.rda")
K12_global_topics <- out

topics_theta <- K12_global_topics$theta

proj <- CRS(' +proj=longlat +ellps=WGS84')
map.extent <- extent(-180,180,-90,90)

newtheme <- rasterTheme(region = colorRampPalette(c("darkseagreen3","orange","red"))( 1000 ))
#newtheme <- colorRampPalette(c("darkseagreen3","orange","red"))( 100 )

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,1]
motif_1_K12 <- matrix(v, nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,2]
motif_2_K12 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,3]
motif_3_K12 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,4]
motif_4_K12 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,5]
motif_5_K12 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,6]
motif_6_K12 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,7]
motif_7_K12 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,8]
motif_8_K12 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,9]
motif_9_K12 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,10]
motif_10_K12 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,11]
motif_11_K12 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,12]
motif_12_K12 <- matrix(v,nrow = 180, ncol = 360)

xmin <- -180
xmax <- 180
ymin <- -90
ymax <- 90

motif_K12_maps <- list()

motif_K12_maps[[1]] <- raster(motif_1_K12, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K12_maps[[2]] <- raster(motif_2_K12, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K12_maps[[3]] <- raster(motif_3_K12, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K12_maps[[4]] <- raster(motif_4_K12, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K12_maps[[5]] <- raster(motif_5_K12, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K12_maps[[6]] <- raster(motif_6_K12, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K12_maps[[7]] <- raster(motif_7_K12, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K12_maps[[8]] <- raster(motif_8_K12, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K12_maps[[9]] <- raster(motif_9_K12, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K12_maps[[10]] <- raster(motif_10_K12, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K12_maps[[11]] <- raster(motif_11_K12, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K12_maps[[12]] <- raster(motif_12_K12, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)


e <- extent(-180, 180, -60, 90)

motif.K12.theta.maps <- list()
for (i in 1:length(motif_K12_maps)){
  motif.K12.theta.maps[[i]] <- levelplot(crop(motif_K12_maps[[i]],e),
                                        main=paste("Map cell contributions to Motif ",i, ", K = 12"),
                                        par.settings=newtheme, 
                                        contour=F,margin=F,
                                        colorkey=list(at=seq(0, 100, 1),
                                                      labels=list(at=c(0,100),
                                                                  labels=c("Low", "High"))))+
    latticeExtra::layer(sp.lines(coast, col='black', lwd=0.5))+
    latticeExtra::layer(sp.polygons(ocean, fill='grey', col='black'))
}

motif.K12.theta.maps[[1]]
motif.K12.theta.maps[[2]]
motif.K12.theta.maps[[3]]
motif.K12.theta.maps[[4]]
motif.K12.theta.maps[[5]]
motif.K12.theta.maps[[6]]
motif.K12.theta.maps[[7]]
motif.K12.theta.maps[[8]]
motif.K12.theta.maps[[9]]
motif.K12.theta.maps[[10]]
motif.K12.theta.maps[[11]]
motif.K12.theta.maps[[12]]
```

## K = 15

Results for K = 15:
```{r, echo=FALSE}

######## k15
load(file="birds_cluster_15.rda")
K15_global_topics <- out

topics_theta <- K15_global_topics$theta

proj <- CRS(' +proj=longlat +ellps=WGS84')
map.extent <- extent(-180,180,-90,90)

newtheme <- rasterTheme(region = colorRampPalette(c("darkseagreen3","orange","red"))( 1000 ))
#newtheme <- colorRampPalette(c("darkseagreen3","orange","red"))( 100 )

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,1]
motif_1_K15 <- matrix(v, nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,2]
motif_2_K15 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,3]
motif_3_K15 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,4]
motif_4_K15 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,5]
motif_5_K15 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,6]
motif_6_K15 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,7]
motif_7_K15 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,8]
motif_8_K15 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,9]
motif_9_K15 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,10]
motif_10_K15 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,11]
motif_11_K15 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,12]
motif_12_K15 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,13]
motif_13_K15 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,14]
motif_14_K15 <- matrix(v,nrow = 180, ncol = 360)

v <- rep(0,64800)
v[as.numeric(rownames(topics_theta))] = topics_theta[,15]
motif_15_K15 <- matrix(v,nrow = 180, ncol = 360)

xmin <- -180
xmax <- 180
ymin <- -90
ymax <- 90

motif_K15_maps <- list()

motif_K15_maps[[1]] <- raster(motif_1_K15, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K15_maps[[2]] <- raster(motif_2_K15, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K15_maps[[3]] <- raster(motif_3_K15, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K15_maps[[4]] <- raster(motif_4_K15, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K15_maps[[5]] <- raster(motif_5_K15, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K15_maps[[6]] <- raster(motif_6_K15, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K15_maps[[7]] <- raster(motif_7_K15, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K15_maps[[8]] <- raster(motif_8_K15, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K15_maps[[9]] <- raster(motif_9_K15, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K15_maps[[10]] <- raster(motif_10_K15, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K15_maps[[11]] <- raster(motif_11_K15, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K15_maps[[12]] <- raster(motif_12_K15, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K15_maps[[13]] <- raster(motif_13_K15, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K15_maps[[14]] <- raster(motif_14_K15, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)
motif_K15_maps[[15]] <- raster(motif_15_K15, xmn=xmin, xmx=xmax, ymn=ymin, ymx=ymax)


e <- extent(-180, 180, -60, 90)

motif.K15.theta.maps <- list()
for (i in 1:length(motif_K15_maps)){
  motif.K15.theta.maps[[i]] <- levelplot(crop(motif_K15_maps[[i]],e),
                                         main=paste("Map cell contributions to Motif ",i, ", K = 15"),
                                         par.settings=newtheme, 
                                         contour=F,margin=F,
                                         colorkey=list(at=seq(0, 100, 1),
                                                       labels=list(at=c(0,100),
                                                                   labels=c("Low", "High"))))+
    latticeExtra::layer(sp.lines(coast, col='black', lwd=0.5))+
    latticeExtra::layer(sp.polygons(ocean, fill='grey', col='black'))
}

motif.K15.theta.maps[[1]]
motif.K15.theta.maps[[2]]
motif.K15.theta.maps[[3]]
motif.K15.theta.maps[[4]]
motif.K15.theta.maps[[5]]
motif.K15.theta.maps[[6]]
motif.K15.theta.maps[[7]]
motif.K15.theta.maps[[8]]
motif.K15.theta.maps[[9]]
motif.K15.theta.maps[[10]]
motif.K15.theta.maps[[11]]
motif.K15.theta.maps[[12]]
motif.K15.theta.maps[[13]]
motif.K15.theta.maps[[14]]
motif.K15.theta.maps[[15]]

```


