---
title: "methClust in Africa"
author: "Kushal K Dey"
date: "12/20/2017"
output: html_document
---

We run **methClust** on the birds presence absence data in the African continent.

```{r setup, include=FALSE}
library(ecostructure)
library(methClust)
library(Biobase)
library(rasterVis)
library(gtools)
library(bigalgebra)
library(sp)
library(rgdal)
library(ggplot2)
library(maps)
library(mapdata)
library(mapplots)
library(scales)
library(ggthemes)

```

## Plotting Maps Function

```{r}
latlong <- get(load("../data/LatLongCells_frame.rda"))
```

```{r}
world_map <- map_data("world")
world_map <- world_map[world_map$region != "Antarctica",] # intercourse antarctica

world_map <- world_map[world_map$long > -18 & world_map$long < 52, ]
world_map <- world_map[world_map$lat > -50 & world_map$lat < 35, ]


p <- ggplot() + coord_fixed() +
  xlab("") + ylab("")

#Add map to base plot
base_world_messy <- p + geom_polygon(data=world_map, aes(x=long, y=lat, group=group), colour="light green", fill="light green")

cleanup <- 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_rect(fill = 'white', colour = 'white'), 
        axis.line = element_line(colour = "white"), legend.position="none",
        axis.ticks=element_blank(), axis.text.x=element_blank(),
        axis.text.y=element_blank())

base_world <- base_world_messy + cleanup

base_world

```

## Extracting Africa presence absence data

```{r}
latlong <- get(load("../data/LatLongCells_frame.rda"))
```

```{r}
idx1 <- which(latlong[,2] > -50 & latlong[,2] < 35)
idx2 <- which(latlong[,1] > -18 & latlong[,1] < 52)
idx <- intersect(idx1, idx2)
length(idx)
latlong2 <- latlong[idx,]
```


```{r}
birds_pa_data <- readRDS("../data/birds_presab_land_breeding_counts.rds")
birds_pa_data_2 <- birds_pa_data[idx, ]
birds_pa_data_3 <- birds_pa_data_2[, which(colSums(birds_pa_data_2)!=0)]
```

```{r}
PlotAssemblageIdx <- function(idx){
  dat <- cbind.data.frame(latlong2, birds_pa_data_3[,idx])
colnames(dat) <- c("Latitude", "Longitude", "Value")
map_data_coloured <- 
  base_world +
  geom_point(data=dat, 
             aes(x=Latitude, y=Longitude, colour=Value), size=0.5) +
  scale_colour_gradient(low = "white", high = "black") 

map_data_coloured
}

```

## Example presence-absence data of bird species 

```{r}
par(mfrow = c(6,2))
PlotAssemblageIdx(1)
PlotAssemblageIdx(28)
PlotAssemblageIdx(100)
PlotAssemblageIdx(546)
PlotAssemblageIdx(176)
PlotAssemblageIdx(50)
PlotAssemblageIdx(24)
PlotAssemblageIdx(1200)
PlotAssemblageIdx(1300)
PlotAssemblageIdx(1400)
PlotAssemblageIdx(1500)
PlotAssemblageIdx(1600)
```

## Application of methClust and visualization

We apply **methClust** on the presence-absence data of relevant bird species along the Wallace Line.

```{r echo=TRUE, eval = FALSE}
topics <- meth_topics(birds_pa_data_3, 1 - birds_pa_data_3, 
            K=10, tol = 10, use_squarem = FALSE)
save(topics, file = "../output/Africa/methClust_10.rda")
```


```{r}
PlotAssemblageIdx2 <- function(dat){
  dat <- cbind.data.frame(latlong2, dat)
colnames(dat) <- c("Latitude", "Longitude", "Value")
map_data_coloured <- 
  base_world +
  geom_point(data=dat, 
             aes(x=Latitude, y=Longitude, colour=Value), size=0.5) +
  scale_colour_gradient(low = "white", high = "black") 

map_data_coloured
}

```

## geo Structure plot representation

### K = 2

```{r}
topics <- get(load("../output/Africa/methClust_2.rda"))
```

```{r, message = FALSE, warning = FALSE}

png(filename="../output/Africa/geostructure_birds_2.png",width = 1000, height = 800)
map("worldHires",
    ylim=c(-50,35), xlim=c(-18,52), # Re-defines the latitude and longitude range
    col = "gray", fill=TRUE, mar=c(0.1,0.1,0.1,0.1))
lapply(1:dim(topics$omega)[1], function(r) 
  add.pie(z=as.integer(100*topics$omega[r,]), 
          x=latlong2[r,1], y=latlong2[r,2], labels=c("","",""),
          radius = 0.5,
          col=c(alpha(2,0.6),alpha(3,0.6))));
dev.off()
```


### K = 3

```{r}
topics <- get(load("../output/Africa/methClust_3.rda"))
```

```{r, message = FALSE, warning = FALSE}

png(filename="../output/Africa/geostructure_birds_3.png",width = 1000, height = 800)
map("worldHires",
    ylim=c(-50,35), xlim=c(-18,52), # Re-defines the latitude and longitude range
    col = "gray", fill=TRUE, mar=c(0.1,0.1,0.1,0.1))
lapply(1:dim(topics$omega)[1], function(r) 
  add.pie(z=as.integer(100*topics$omega[r,]), 
          x=latlong2[r,1], y=latlong2[r,2], labels=c("","",""),
          radius = 0.5,
          col=c(alpha(2,0.6),alpha(3,0.6), alpha(4,0.6))));
dev.off()
```

### K = 4

```{r}
topics <- get(load("../output/Africa/methClust_4.rda"))
```

```{r, message = FALSE, warning = FALSE}

png(filename="../output/Africa/geostructure_birds_4.png",width = 1000, height = 800)
map("worldHires",
    ylim=c(-50,35), xlim=c(-18,52), # Re-defines the latitude and longitude range
    col = "gray", fill=TRUE, mar=c(0.1,0.1,0.1,0.1))
lapply(1:dim(topics$omega)[1], function(r) 
  add.pie(z=as.integer(100*topics$omega[r,]), 
          x=latlong2[r,1], y=latlong2[r,2], labels=c("","",""),
          radius = 0.5,
          col=c(alpha(2,0.6),alpha(3,0.6), alpha(4,0.6), alpha(5,0.6))));
dev.off()
```

### K = 5

```{r}
topics <- get(load("../output/Africa/methClust_5.rda"))
```

```{r, message = FALSE, warning = FALSE}

png(filename="../output/Africa/geostructure_birds_5.png",width = 1000, height = 800)
map("worldHires",
    ylim=c(-50,35), xlim=c(-18,52), # Re-defines the latitude and longitude range
    col = "gray", fill=TRUE, mar=c(0.1,0.1,0.1,0.1))
lapply(1:dim(topics$omega)[1], function(r) 
  add.pie(z=as.integer(100*topics$omega[r,]), 
          x=latlong2[r,1], y=latlong2[r,2], labels=c("","",""),
          radius = 0.5,
          col=c(alpha(2,0.6),alpha(3,0.6), alpha(4,0.6), alpha(5,0.6), alpha(6,0.6))));
dev.off()
```
