---
title: 'methClust applications : Himalayan bird species'
author: "Kushal K Dey"
date: "12/12/2017"
output: html_document
---

```{r message = FALSE, warning=FALSE}
library(ecostructure)
library(methClust)
library(slam)
```

We use **methClust** to fit the presence-absence matrix of regional assemblage fields.

## With Abundance 

### Data Processing

We take the counts data of abundances as Binomial counts $X_j$ with the assumption.

$$  X_j \sim Bin (X_j, p_j)  \hspace{1 cm}  X_j > 0$$
and

$$ X_j \sim Bin (1, p_j)  \hspace{1 cm} X_j  =0 $$

Then we fit grades of memberships on $p_j$ values. 

Note that this model is definitely not a good model, but it does take into account the abundances instead of presence and absences, and I am trying to see how that affects the Grades of memberships.

```{r}
reg_data <- readRDS("../data/Him_site_x_cell_matrix.rds")
#reg_data[reg_data > 0] = 1
meth <- reg_data
reg_data_2 <- reg_data
reg_data_2[reg_data_2 > 0] = 1
unmeth <- 1 - reg_data_2
```


### Fitting methClust model (different Ks)


```{r echo=TRUE, eval=FALSE}

topics <- meth_topics(meth, unmeth, 
                      K=4, tol = 10, use_squarem = FALSE)

```

### Visualization of methClust model

We plot the Structure plot representation of the regional assemblage grades of membership for different values of $K=2,3,4, 5$.

#### K=2 

```{r}
topics <- get(load("../output/methClust_Himalayas_K_2_abundance.rda"))
data <- get(load(system.file("extdata", "HimalayanBirdsData.rda",
                             package = "ecostructure")))
grid_metadata <- pData(phenoData(data))
head(grid_metadata)
elevation_metadata=grid_metadata$Elevation;
east_west_dir = grid_metadata$WorE;
omega <- topics$omega
BlockStructure(omega, blocker_metadata = east_west_dir,
               order_metadata = elevation_metadata,
               yaxis_label = "Elevation",
               levels_decreasing = FALSE)
```

#### K=3

```{r}
topics <- get(load("../output/methClust_Himalayas_K_3_abundance.rda"))
omega <- topics$omega
BlockStructure(omega, blocker_metadata = east_west_dir,
               order_metadata = elevation_metadata,
               yaxis_label = "Elevation",
               levels_decreasing = FALSE)
```

#### K=4

```{r}
topics <- get(load("../output/methClust_Himalayas_K_4_abundance.rda"))
omega <- topics$omega
BlockStructure(omega, blocker_metadata = east_west_dir,
               order_metadata = elevation_metadata,
               yaxis_label = "Elevation",
               levels_decreasing = FALSE)
```


## Presence absence matrix

Now we consider a more relaistic model of presence absence matrix, where we first concvert positive counts to just 1 

$$ Y_j = I (X_j > 0) $$

and then assume 

$$  Y_j \sim Ber(p_j) $$

and then fit grades of membership model on the $p_j$ values. 


```{r}
reg_data <- readRDS("../data/Him_site_x_cell_matrix.rds")
reg_data[reg_data > 0] = 1
meth <- reg_data
unmeth <- 1 - reg_data
```


### Fitting methClust model (different Ks)


```{r echo=TRUE, eval=FALSE}
topics <- meth_topics(meth, unmeth, 
                      K=4, tol = 10, use_squarem = FALSE)

```

### Visualization of methClust model

We plot the Structure plot representation of the regional assemblage grades of membership for different values of $K=2,3,4, 5$.

#### K=2 

```{r}
topics <- get(load("../output/methClust_Himalayas_K_2.rda"))
data <- get(load(system.file("extdata", "HimalayanBirdsData.rda",
                             package = "ecostructure")))
grid_metadata <- pData(phenoData(data))
head(grid_metadata)
elevation_metadata=grid_metadata$Elevation;
east_west_dir = grid_metadata$WorE;
omega <- topics$omega
BlockStructure(omega, blocker_metadata = east_west_dir,
               order_metadata = elevation_metadata,
               yaxis_label = "Elevation",
               levels_decreasing = FALSE)
```

#### K=3

```{r}
topics <- get(load("../output/methClust_Himalayas_K_3.rda"))
omega <- topics$omega
BlockStructure(omega, blocker_metadata = east_west_dir,
               order_metadata = elevation_metadata,
               yaxis_label = "Elevation",
               levels_decreasing = FALSE)
```

#### K=4

```{r}
topics <- get(load("../output/methClust_Himalayas_K_4.rda"))
omega <- topics$omega
BlockStructure(omega, blocker_metadata = east_west_dir,
               order_metadata = elevation_metadata,
               yaxis_label = "Elevation",
               levels_decreasing = FALSE)
```



