---
title: "methClust along Wallace Line"
author: "Kushal K Dey"
date: "12/18/2017"
output: html_document
---

We apply methClust to cells along the Wallace line. We wish to check how the Wallace Line affects the communities of bird species. 

```{r setup, include=FALSE}
library(ecostructure)
library(methClust)
library(Biobase)
library(rasterVis)
library(gtools)
library(bigalgebra)
library(sp)
library(rgdal)
library(ggplot2)
library(maps)
library(mapdata)
library(mapplots)
library(scales)
library(ggthemes)

```

## Plotting Maps Function

```{r}
latlong <- get(load("../data/LatLongCells_frame.rda"))
```

```{r}
world_map <- map_data("world")
world_map <- world_map[world_map$region != "Antarctica",] # intercourse antarctica

world_map <- world_map[world_map$long > 100 & world_map$long < 150, ]
world_map <- world_map[world_map$lat > -10 & world_map$lat < 10, ]


p <- ggplot() + coord_fixed() +
  xlab("") + ylab("")

#Add map to base plot
base_world_messy <- p + geom_polygon(data=world_map, aes(x=long, y=lat, group=group), colour="light green", fill="light green")

cleanup <- 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_rect(fill = 'white', colour = 'white'), 
        axis.line = element_line(colour = "white"), legend.position="none",
        axis.ticks=element_blank(), axis.text.x=element_blank(),
        axis.text.y=element_blank())

base_world <- base_world_messy + cleanup

base_world

```

## Extracting Wallace Line presence absence data

```{r}
latlong <- get(load("../data/LatLongCells_frame.rda"))
```

```{r}
idx1 <- which(latlong[,2] > -10 & latlong[,2] < 10)
idx2 <- which(latlong[,1] > 100 & latlong[,1] < 150)
idx <- intersect(idx1, idx2)
length(idx)
latlong2 <- latlong[idx,]
```

```{r}
birds_pa_data <- readRDS("../data/birds_presab_land_breeding_counts.rds")
birds_pa_data_2 <- birds_pa_data[idx, ]
birds_pa_data_3 <- birds_pa_data_2[, which(colSums(birds_pa_data_2)!=0)]
```

```{r}
PlotAssemblageIdx <- function(idx){
  dat <- cbind.data.frame(latlong2, birds_pa_data_3[,idx])
colnames(dat) <- c("Latitude", "Longitude", "Value")
map_data_coloured <- 
  base_world +
  geom_point(data=dat, 
             aes(x=Latitude, y=Longitude, colour=Value), size=0.5) +
  scale_colour_gradient(low = "white", high = "black") 

map_data_coloured
}

```

## Example presence-absence data of bird species 

```{r}
par(mfrow = c(6,2))
PlotAssemblageIdx(1)
PlotAssemblageIdx(10)
PlotAssemblageIdx(100)
PlotAssemblageIdx(500)
PlotAssemblageIdx(700)
PlotAssemblageIdx(1000)
PlotAssemblageIdx(1100)
PlotAssemblageIdx(1200)
PlotAssemblageIdx(1300)
PlotAssemblageIdx(1400)
PlotAssemblageIdx(1500)
PlotAssemblageIdx(1600)
```

## Application of methClust and visualization

We apply **methClust** on the presence-absence data of relevant bird species along the Wallace Line.

```{r echo=TRUE, eval = FALSE}
topics <- meth_topics(birds_pa_data_3, 1 - birds_pa_data_3, 
            K=5, tol = 0.1, use_squarem = FALSE)
```

```{r}
topics <- get(load("../output/methClust_wallace_K_2.rda"))
```

```{r}
PlotAssemblageIdx2 <- function(dat){
  dat <- cbind.data.frame(latlong2, dat)
colnames(dat) <- c("Latitude", "Longitude", "Value")
map_data_coloured <- 
  base_world +
  geom_point(data=dat, 
             aes(x=Latitude, y=Longitude, colour=Value), size=0.5) +
  scale_colour_gradient(low = "white", high = "black") 

map_data_coloured
}

```

```{r}
PlotAssemblageIdx2(topics$omega[,1])
PlotAssemblageIdx2(topics$omega[,2])
```

You can see that both clusters have partial grades of membership along the Wallace line (points shaded grey). But one cluster is mainly driven by bird species to the north of the Wallace Line  while the other cluster driven by bird species to the south of Wallace Line. 

## geo Structure plot representation

### K = 2

```{r, message = FALSE, warning = FALSE}
png(filename="../output/geostructure_birds.png",width = 1000, height = 800)
map("worldHires",
    ylim=c(-10,10), xlim=c(100,150), # Re-defines the latitude and longitude range
    col = "gray", fill=TRUE, mar=c(0.1,0.1,0.1,0.1))
lapply(1:dim(topics$omega)[1], function(r) 
  add.pie(z=as.integer(100*topics$omega[r,]), 
          x=latlong2[r,1], y=latlong2[r,2], labels=c("","",""),
          radius = 0.5,
          col=c(alpha(2,0.6),alpha(3,0.6))));
dev.off()
```

### K = 3

```{r}
topics <- get(load("../output/methClust_wallace_K_3.rda"))
```

```{r, message = FALSE, warning = FALSE}
png(filename="../output/geostructure_birds_3.png",width = 1000, height = 800)
map("worldHires",
    ylim=c(-10,10), xlim=c(100,150), # Re-defines the latitude and longitude range
    col = "gray", fill=TRUE, mar=c(0.1,0.1,0.1,0.1))
lapply(1:dim(topics$omega)[1], function(r) 
  add.pie(z=as.integer(100*topics$omega[r,]), 
          x=latlong2[r,1], y=latlong2[r,2], labels=c("","",""),
          radius = 0.5,
          col=c(alpha(2,0.6),alpha(3,0.6), alpha(4,0.6))));
dev.off()
```

### K = 4

```{r}
topics <- get(load("../output/methClust_wallace_K_4.rda"))
```

```{r, message = FALSE, warning = FALSE}
png(filename="../output/geostructure_birds_4.png",width = 1000, height = 800)
map("worldHires",
    ylim=c(-10,10), xlim=c(100,150), # Re-defines the latitude and longitude range
    col = "gray", fill=TRUE, mar=c(0.1,0.1,0.1,0.1))
lapply(1:dim(topics$omega)[1], function(r) 
  add.pie(z=as.integer(100*topics$omega[r,]), 
          x=latlong2[r,1], y=latlong2[r,2], labels=c("","",""),
          radius = 0.5,
          col=c(alpha(2,0.6),alpha(3,0.6), alpha(4,0.6), alpha(5,0.6))));
dev.off()
```

### K = 5

```{r}
topics <- get(load("../output/methClust_wallace_K_5.rda"))
```

```{r, message = FALSE, warning = FALSE}
png(filename="../output/geostructure_birds_5.png",width = 1000, height = 800)
map("worldHires",
    ylim=c(-10,10), xlim=c(100,150), # Re-defines the latitude and longitude range
    col = "gray", fill=TRUE, mar=c(0.1,0.1,0.1,0.1))
lapply(1:dim(topics$omega)[1], function(r) 
  add.pie(z=as.integer(100*topics$omega[r,]), 
          x=latlong2[r,1], y=latlong2[r,2], labels=c("","",""),
          radius = 0.5,
          col=c(alpha(2,0.6),alpha(3,0.6), alpha(4,0.6), alpha(5,0.6), alpha(6,0.6))));
dev.off()
```

