---
title: 'Simulation example : methClust'
author: "Kushal K Dey"
date: "11/29/2017"
output: 
  html_document:
    theme : cosmo
    code_folding: hide
---


```{r}
library(methClust)
```

# Example

Here we present a small demo of how **methClust** can be used to fit a Grade of Membership model on a simulated example. This example will be replaced with a more appropriate real example with the further development of this package.

We first define the grades of membership matrix $\omega_{N \times K}$ where $N$ is the number of samples and $K$ is the number of clusters

```{r eval=TRUE, warning=FALSE, fig.align="center", fig.show="asis"}
n.out <- 500
omega_sim <- rbind( cbind( rep(1, n.out), rep(0, n.out)),
                    cbind( rep(0, n.out), rep(1, n.out)),
                    cbind( seq(0.6, 0.4, length.out = n.out),
                           1- seq(0.6, 0.4,length.out=n.out)) )

K <- dim(omega_sim)[2]
barplot(t(omega_sim),
        col = 2:(K+1),
        axisnames = F, space = 0,
        border = NA,
        main=paste("No. of clusters=", K),
        las=1, ylim = c(0,1), cex.axis=1.5,cex.main=1.4)

```


Next we define the cluster profiles $f_{K \times B}$ where $0 \leq f_{kb} \leq 1$ and $B$ denotes the number of genomic bins that the genome has been partitioned into.

```{r}
m.out <- 200
freq_sim <- cbind(c(rep(0.8, m.out), rep(0.2, m.out), rep(0.5, m.out), rep(0.01, m.out)),
                  c(rep(0.01, m.out), rep(0.01, m.out), rep(0.5, m.out), rep(0.8, m.out)))

```

We next define matrices for the number of methylated sites M and unmethylated sites U per sample annd per bin, which are assumed to obey Equation 1.

```{r}
prob <- omega_sim %*% t(freq_sim)
Y <- matrix(rpois(dim(prob)[1]*dim(prob)[2], 1000), dim(prob)[1], dim(prob)[2])

M <- matrix(0, dim(Y)[1], dim(Y)[2])

for(m in 1:dim(Y)[1]){
  for(n in 1:dim(Y)[2]){
    M[m,n] <- rbinom(1, Y[m,n], prob = prob[m,n])
  }
}

U = Y - M

```

Check the dimensions the methylation matrix M and the unmethylation matrix (they should be same).

```{r}
dim(M)
dim(U)
```

Now we fit **methClust**

```{r echo=TRUE, eval=TRUE}
topics <- meth_topics(M, U, K=2, tol = 10, use_squarem = FALSE)
```

The esimated bar chart representation of the grades of membership are as follows 

```{r eval=TRUE, warning=FALSE, fig.align="center", fig.show="asis"}
barplot(t(topics$omega),
        col = 2:(K+1),
        axisnames = F, space = 0,
        border = NA,
        main=paste("No. of clusters=", K),
        las=1, ylim = c(0,1), cex.axis=1.5,cex.main=1.4)
```

