---
title: "methClust in Europe"
author: "Kushal K Dey"
date: "12/25/2017"
output: html_document
---

We apply methClust to cells in Europe. We wish to check how the Wallace Line affects the communities of bird species. 

```{r setup, include=FALSE}
library(ecostructure)
library(methClust)
library(Biobase)
library(rasterVis)
library(gtools)
library(bigalgebra)
library(sp)
library(rgdal)
library(ggplot2)
library(maps)
library(mapdata)
library(mapplots)
library(scales)
library(ggthemes)
```


## Plotting Maps Function

```{r}
latlong <- get(load("../data/LatLongCells_frame.rda"))
```

```{r}
world_map <- map_data("world")
world_map <- world_map[world_map$region != "Antarctica",] # intercourse antarctica

world_map <- world_map[world_map$long > -15 & world_map$long < 50, ]
world_map <- world_map[world_map$lat > 36 & world_map$lat < 70, ]


p <- ggplot() + coord_fixed() +
  xlab("") + ylab("")

#Add map to base plot
base_world_messy <- p + geom_polygon(data=world_map, aes(x=long, y=lat, group=group), colour="light green", fill="light green")

cleanup <- 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background = element_rect(fill = 'white', colour = 'white'), 
        axis.line = element_line(colour = "white"), legend.position="none",
        axis.ticks=element_blank(), axis.text.x=element_blank(),
        axis.text.y=element_blank())

base_world <- base_world_messy + cleanup

base_world

```

## Extracting Europe birds presence absence data

```{r}
latlong <- get(load("../data/LatLongCells_frame.rda"))
```

```{r}
idx1 <- which(latlong[,2] > 36 & latlong[,2] < 70)
idx2 <- which(latlong[,1] > -15 & latlong[,1] < 50)
idx <- intersect(idx1, idx2)
length(idx)
latlong2 <- latlong[idx,]
```

```{r}
birds_pa_data <- readRDS("../data/birds_presab_land_breeding_counts.rds")
birds_pa_data_2 <- birds_pa_data[idx, ]
birds_pa_data_3 <- birds_pa_data_2[, which(colSums(birds_pa_data_2) > 10)]
```

## Example presence-absence data of bird species 

```{r}
par(mfrow = c(3,2))
PlotAssemblageIdx(1)
PlotAssemblageIdx(10)
PlotAssemblageIdx(100)
PlotAssemblageIdx(200)
PlotAssemblageIdx(300)
PlotAssemblageIdx(400)
```



## Application of methClust and visualization

We apply **methClust** on the presence-absence data of relevant bird species in Europe. 

```{r echo=TRUE, eval = FALSE}
topics <- meth_topics(birds_pa_data_3, 1 - birds_pa_data_3, 
            K=7, tol = 0.1, use_squarem = FALSE)
save(topics, file = "../output/Europe/methClust_7.rda")
```

